---
title: "Car Transmision and Miles Per Gallon V2"
output: html_document
---
Motor Trend, a magazine about the automobile industry
Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

  “Is an automatic or manual transmission better for MPG”
  "Quantify the MPG difference between automatic and manual transmissions"


```{r}
library(ggplot2)
library(caret)
data(mtcars)
summary(mtcars)
#mtcars$am = ifelse(mtcars$am = 0, 'manual', )
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(mtcars$mpg ~ ., data = mtcars)
plot(factor(mtcars$am), mtcars$mpg,data=mtcars)
```

```{r}
fit = lm(mpg ~ ., data=mtcars)
coef(fit)


```

```{r}
qq <- qplot(am+ disp,mpg, colour = gear, data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(hp, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(cyl, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(disp, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(drat, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(wt, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(qsec, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(vs, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(gear, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

qq <- qplot(carb, mpg, colour = factor(am), data=mtcars)
qq + geom_smooth(method='lm', formula = y~x)

```

```{r}
featurePlot(x=mtcars[,c("cyl", "disp")], y = mtcars$mpg, plot = "pairs")

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.   

Appendix

